Let's talk about Thurstone \& Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2025-05-29 _tex/paper2.tex
Built site for gh-pages

Fortunately, SCMs and DAGs support identification analysis through two
key advantages\footnote{These topics are beyond the scope of this study,
  thus, readers seeking a more profound understanding can refer to
  introductory papers such as \citet{Pearl_2010}, \citet{Rohrer_2018},
  \citet{Pearl_2019}, and \citet{Cinelli_et_al_2020}, and introductory
  books like \citet{Pearl_et_al_2018}, \citet{Neal_2020}, and
  \citet{McElreath_2020} are useful. For more advanced study, seminal
  papers such as \citet{Neyman_et_al_1923}, \citet{Rubin_1974},
  \citet{Spirtes_et_al_1991}, and \citet{Sekhon_2009}, along with books
  such as \citet{Pearl_2009}, \citet{Morgan_et_al_2014}, and
  \citet{Hernan_et_al_2020}, are recommended.}. First, regardless of
complexity, they can represent various causal structures using only five
fundamental building blocks \citep{Neal_2020, McElreath_2024}. This
feature allows researchers to decompose complex structures into
manageable components, facilitating their analysis. Second, they depict
causal relationships in a non-parametric way. This flexibility enables
feasible identification strategies without requiring specification of
the types of variables, the functional forms relating them, or the
parameters of those functional forms \citep{Pearl_et_al_2016}.

@@

Here, each element of \(I\) represents a unique individual \(i\) or
\(h\), where \(n_{I}\) denotes the total number of individuals.
Similarly, each element of \(J\) corresponds to a unique judge \(j\),
with \(n_{J}\) indicating the total number of judges. Moreover, each row
of \(IA\) represents a unique pairing of individuals \(i, h\) with
stimuli \(a, b\). As a result, the matrix \(IA\) contains
\(n_{I} \cdot n_{A}\) rows and \(2\) columns, where \(n_{A}\) specifies
the number of stimuli available per individual. Likewise, each row of
\(JK\) associates a judge \(j\) with a (repeated) judgment index \(k\).
Consequently, the matrix \(JK\) has \(n_{J} \cdot n_{K}\) rows and \(2\)
columns, where \(n_{K}\) indicates the number of repeated judgments each
judge makes.

Additionally, we construct the matrix \(R\) to map each row of the
\(IA\) matrix with a corresponding row from the \(JK\) matrix. This
matrix has \(n\) rows and \(6\) columns, where
\(n = {n_{I} \cdot n_{A} \choose 2} \cdot n_{J} \cdot n_{K}\). Here, the
term \({n_{I} \cdot n_{A} \choose 2}\) represents the binomial
coefficient, which quantifies the total number of unique comparisons
possible between every pair of stimuli generated by each pair of
individuals in the population. Thus, we define the matrix as follows:

\begin{equation}\phantomsection\label{eq-mat12}{
R = \begin{bmatrix}
1 & 1 & 1 & 2 & 1 & 1 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
1 & 1 & 1 & 2 & 1 & n_{K} \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
i & a & h & b & j & k \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
n_{I} & n_{A}-1 & n_{I} & n_{A} & n_{J} & 1 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
n_{I} & n_{A}-1 & n_{I} & n_{A} & n_{J} & n_{K}
\end{bmatrix}
}\end{equation}

It is easier to visualize the structure of these vectors and matrices by
considering an example. Assuming \(n_{I} = 5\), \(n_{A} = 2\),
\(n_{J} = 3\), and \(n_{K} = 3\), the vectors and matrices described in
Equations (\ref{eq-mat11}) and (\ref{eq-mat12}) take the following form:
\begin{equation}\phantomsection\label{eq-mat13}{
I = \begin{bmatrix}
1 \\
2 \\
3 \\
4 \\
5 
\end{bmatrix} ; \;
J = \begin{bmatrix}
1 \\
2 \\
3 
\end{bmatrix} ; \;
IA = \begin{bmatrix}
1 & 1 \\
1 & 2 \\
2 & 1 \\
2 & 2 \\
3 & 1 \\
3 & 2 \\
4 & 1 \\
4 & 2 \\
5 & 1 \\
5 & 2 
\end{bmatrix} ; \;
JK = \begin{bmatrix}
1 & 1 \\
1 & 2 \\
1 & 3 \\
2 & 1 \\
2 & 2 \\
2 & 3 \\
3 & 1 \\
3 & 2 \\
3 & 3 
\end{bmatrix} ; \;
R = \begin{bmatrix}
1 & 1 & 1 & 2 & 1 & 1 \\
1 & 1 & 1 & 2 & 1 & 2 \\
1 & 1 & 1 & 2 & 1 & 3 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
1 & 1 & 5 & 2 & 1 & 1 \\
1 & 1 & 5 & 2 & 1 & 2 \\
1 & 1 & 5 & 2 & 1 & 3 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
4 & 2 & 5 & 2 & 3 & 1 \\
4 & 2 & 5 & 2 & 3 & 2 \\
4 & 2 & 5 & 2 & 3 & 3 \\
5 & 1 & 5 & 2 & 3 & 1 \\
5 & 1 & 5 & 2 & 3 & 2 \\
5 & 1 & 5 & 2 & 3 & 3 
\end{bmatrix}
}\end{equation}

@@

Finally, we define the \emph{sample mechanism} \(S\) in Equation
(\ref{eq-mat25}), which maps each element of \(S_{IA}\) to every element
of \(S_{JK}\). Each element \(s_{(i,a,h,b,j,k)}\) is a binary value
indicating the presence or absence of data rows in the matrix \(R\)
resulting from the sample mechanism, as in Equation (\ref{eq-mat26}).
Thus, the vector contains \(n\) elements, matching the number of rows in
\(R\), and the sum of its elements represents the total data sample:
\(n^{s} = \binom{n^{s}_{I} \cdot n^{s}_{A}}{2} \cdot n^{s}_{J} \cdot n^{s}_{K}\).
Here, the term \({n^{s}_{I} \cdot n^{s}_{A} \choose 2}\) represents the
binomial coefficient, which quantifies the total number of unique
comparisons possible between every pair of sampled stimuli generated by
each pair of sampled individuals.

\begin{equation}\phantomsection\label{eq-mat25}{
S = \begin{bmatrix}
s_{(1,1,1,2,1,1)} \\
\vdots \\
s_{(1,1,1,2,1,n_{K})} \\
\vdots \\
s_{(i,a,h,b,j,k)} \\
\vdots \\
s_{(n_{I},n_{A}-1,n_{I},n_{A},n_{J},1)} \\
\vdots \\
s_{(n_{I},n_{A}-1,n_{I},n_{A},n_{J},1)}
\end{bmatrix}
}\end{equation}

@@

Note that while the joint distribution in Equation (\ref{eq-mat51})
includes the probability distributions of the sampling and comparison
mechanisms, \(P(S)\) and \(P(C)\), as well as those of the predictor
variables--\(P(X_{IA})\), \(P(X_{I})\), \(P(Z_{JK})\), and
\(P(Z_{J})\)--we omit all of these probabilities from the statistical
model \ref{fig-cj16_stat}. This omission is justified because, while
these distributions contribute to the overall joint distribution of the
data, the variables \(S\), \(C\), \(X_{IA}\), \(X_{I}\), \(Z_{JK}\), and
\(Z_{J}\) are observed and independent of any other variable in the
model. As observed variables, they do not require distributional
assumptions in the same way the idiosyncratic errors do. Furthermore,
their independence follows from the underlying random selection
procedures that govern the variables\footnote{Randomization ensures that
  data--and, by extension, an estimator--satisfies several key
  identification properties, such as common support, no interference,
  and consistency. The most critical property, however, is the
  elimination of confounding. \emph{Confounding} occurs when an external
  variable, such as \(X_{I}\), simultaneously influences both the
  outcome (e.g., \(O_{R}\)) and a variable of interest (e.g., \(S\)),
  resulting in spurious associations between the latter two
  \citep{Everitt_et_al_2010}. Randomization ensure the absence of
  confounding by effectively decoupling the association between the
  variable of interest and any other variable, except for the outcome
  itself. For a more detailed discussion on the benefits of
  randomization, see \citet{Pearl_2009}, \citet{Morgan_et_al_2014},
  \citet{Neal_2020}, and \citet{Hernan_et_al_2020}.}.

@@

Thurstone introduced the Law of Comparative Judgment to measure
psychological traits of stimuli through pairwise comparisons
\citep{Thurstone_1927a, Thurstone_1927b}. In its general form, the
theory models single-judge comparisons across multiple, potentially
correlated stimuli. Each comparison produces a dichotomous outcome
indicating which stimulus the judge perceives as having a higher trait
level. However, Thurstone identified one key challenge with its
measurement model: it required estimating more ``unknown'' parameters
than the number of available pairwise comparisons
\citep{Thurstone_1927b}. To address this issue and facilitate the
theory's practical applicability, he formulated five cases derived from
this general form, each progressively incorporating several simplifying
assumptions.

Of these five cases, the CJ literature has predominantly relied on Case
V to measure various psychological traits, as shown in studies like
\citet{Kimbell_2012}, \citet{Jones_et_al_2015}, and
\citet{Boonen_et_al_2020}. This preference largely stems from the
widespread adoption of the BTL model, which offers a simplified
statistical formulation of the case. The BTL model mirrors the
assumptions of Case V--namely, equal discriminal dispersions and zero
correlation for the stimuli' discriminal processes--with one notable
distinction. While Case V assumes normally distributed discriminal
processes, the BTL model uses the more mathematically tractable Logistic
distribution \citep{Andrich_1978, Bramley_2008}. Although this
substitution has a minimal impact on trait estimation or model
interpretation \citep{vanderLinden_et_al_2017_I, McElreath_2021},
Thurstone acknowledged that assuming equal discriminal dispersions and
zero correlation among stimuli' discriminal processes could introduce
significant problems. In particular, he recognized that while these
assumptions simplify the trait measurement model, they may not capture
the complexity of some traits or account for heterogeneous stimuli, such
as handwritten texts or English compositions
\citep{Thurstone_1927a, Andrich_1978, Bramley_2008, Kelly_et_al_2022}.
As a result, they can lead to unreliable and inaccurate trait estimates
\citep{Ackerman_1989, Zimmerman_1994, McElreath_2020, Hoyle_et_al_2023}.

Furthermore, because Thurstone aimed primarily to produce a trait's
``coarse scaling'' and allocate the compared stimuli along this
continuum \citep[pp.~269]{Thurstone_1927a}, his theory did not specify
how to use the resulting trait estimates for statistical inference. The
CJ tradition has attempted to address this gap by separating trait
estimation from hypothesis testing. Specifically, CJ studies often rely
on point estimates of the traits--typically the BTL scores or its
transformations--to conduct statistical inference. Although this
separation simplifies CJ data analysis, the statistical literature warns
that this practice can introduce bias into the analysis and compromise
the reliability of statistical inferences
\citep{McElreath_2020, Kline_et_al_2023, Hoyle_et_al_2023}.

Thus, to address the limitations of Thurstone's Case V and the BTL
model--particularly their strong assumptions and the disconnect between
trait measurement and hypothesis testing--this study extends Thurstone's
general form using a systematic, integrated approach that combines
Causal and Bayesian inference methods. The approach begins by
formulating a conceptual model represented by a Structural Causal Model
(SCM) and a Directed Acyclic Graph (DAG). This model integrates
Thurstone's core theoretical principles--such as the discriminal
processes of stimuli--alongside key CJ experimental design features,
including judges' bias, sampling procedures, and comparison mechanisms,
thereby disentangling the causal processes in the CJ system.

@@

@@

This section introduces fundamental statistical and causal inference
concepts necessary for understanding the core theoretical principles
described in this document. It does not, however, offer a comprehensive
overview of causal inference methods. Readers seeking more in-depth
understanding may wish to explore introductory papers such as
\citet{Pearl_2010}, \citet{Rohrer_2018}, \citet{Pearl_2019}, and
\citet{Cinelli_et_al_2020}. They may also find it helpful to consult
introductory books like \citet{Pearl_et_al_2018}, \citet{Neal_2020}, and
\citet{McElreath_2020}. For more advanced study, readers may refer to
seminal intermediate papers such as \citet{Neyman_et_al_1923},
\citet{Rubin_1974}, \citet{Spirtes_et_al_1991}, and \citet{Sekhon_2009},
as well as books such as \citet{Pearl_2009}, \citet{Morgan_et_al_2014},
and \citet{Hernan_et_al_2020}.

>>>>

Fortunately, SCMs and DAGs support identification analysis through two
key advantages\footnote{These topics are beyond the scope of this study,
  thus, readers seeking a more profound understanding can refer to
  introductory papers such as \citet{Pearl_2010}, \citet{Rohrer_2018},
  \citet{Pearl_2019}, and \citet{Cinelli_et_al_2020}, and introductory
  books like \citet{Pearl_et_al_2018}, \citet{Neal_2020}, and
  \citet{McElreath_2020} are useful. For more advanced study, seminal
  papers such as \citet{Neyman_et_al_1923}, \citet{Rubin_1974},
  \citet{Spirtes_et_al_1991}, and \citet{Sekhon_2009}, along with books
  such as \citet{Pearl_2009}, \citet{Morgan_et_al_2014}, and
  \citet{Hernan_et_al_2025}, are recommended.}. First, regardless of
complexity, they can represent various causal structures using only five
fundamental building blocks \citep{Neal_2020, McElreath_2024}. This
feature allows researchers to decompose complex structures into
manageable components, facilitating their analysis. Second, they depict
causal relationships in a non-parametric way. This flexibility enables
feasible identification strategies without requiring specification of
the types of variables, the functional forms relating them, or the
parameters of those functional forms \citep{Pearl_et_al_2016}.

@@

Here, each element of \(I\) represents a unique individual \(i\) or
\(h\), where \(n_{I}\) denotes the total number of individuals.
Similarly, each element of \(J\) corresponds to a unique judge \(j\),
with \(n_{J}\) indicating the total number of judges. Moreover, each row
of \(IA\) represents a unique pairing of individuals \(i, h\) with
stimuli \(a, b\). As a result, the matrix \(IA\) contains
\(n_{I} \cdot n_{A}\) rows and \(2\) columns, where \(n_{A}\) specifies
the number of stimuli available per individual. Likewise, each row of
\(JK\) associates a judge \(j\) with a (repeated) judgment index \(k\).
Consequently, the matrix \(JK\) has \(n_{J} \cdot n_{K}\) rows and \(2\)
columns, where \(n_{K}\) indicates the number of repeated judgments each
judge makes. \begin{equation}\phantomsection\label{eq-mat12}{
R = \begin{bmatrix}
1 & 1 & 1 & 2 & 1 & 1 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
1 & 1 & 1 & 2 & 1 & n_{K} \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
i & a & h & b & j & k \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
n_{I} & n_{A}-1 & n_{I} & n_{A} & n_{J} & 1 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
n_{I} & n_{A}-1 & n_{I} & n_{A} & n_{J} & n_{K}
\end{bmatrix}
}\end{equation}

Additionally, we construct the matrix \(R\) to map each row of the
\(IA\) matrix with a corresponding row from the \(JK\) matrix. This
matrix has \(n\) rows and \(6\) columns, where
\(n = {n_{I} \cdot n_{A} \choose 2} \cdot n_{J} \cdot n_{K}\). Here, the
term \({n_{I} \cdot n_{A} \choose 2}\) represents the binomial
coefficient, which quantifies the total number of unique comparisons
possible between every pair of stimuli generated by each pair of
individuals in the population. Thus, we define the matrix as in Equation
(\ref{eq-mat12}).

It is easier to visualize the structure of the previously defined
vectors and matrices by considering an example. Assuming \(n_{I} = 5\),
\(n_{A} = 2\), \(n_{J} = 3\), and \(n_{K} = 3\), the vectors and
matrices described in Equations (\ref{eq-mat11}) and (\ref{eq-mat12})
take the form as in Equation (\ref{eq-mat13}).
\begin{equation}\phantomsection\label{eq-mat13}{
I = \begin{bmatrix}
1 \\
2 \\
3 \\
4 \\
5 
\end{bmatrix} ; \;
J = \begin{bmatrix}
1 \\
2 \\
3 
\end{bmatrix} ; \;
IA = \begin{bmatrix}
1 & 1 \\
1 & 2 \\
2 & 1 \\
2 & 2 \\
3 & 1 \\
3 & 2 \\
4 & 1 \\
4 & 2 \\
5 & 1 \\
5 & 2 
\end{bmatrix} ; \;
JK = \begin{bmatrix}
1 & 1 \\
1 & 2 \\
1 & 3 \\
2 & 1 \\
2 & 2 \\
2 & 3 \\
3 & 1 \\
3 & 2 \\
3 & 3 
\end{bmatrix} ; \;
R = \begin{bmatrix}
1 & 1 & 1 & 2 & 1 & 1 \\
1 & 1 & 1 & 2 & 1 & 2 \\
1 & 1 & 1 & 2 & 1 & 3 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
1 & 1 & 5 & 2 & 1 & 1 \\
1 & 1 & 5 & 2 & 1 & 2 \\
1 & 1 & 5 & 2 & 1 & 3 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
4 & 2 & 5 & 2 & 3 & 1 \\
4 & 2 & 5 & 2 & 3 & 2 \\
4 & 2 & 5 & 2 & 3 & 3 \\
5 & 1 & 5 & 2 & 3 & 1 \\
5 & 1 & 5 & 2 & 3 & 2 \\
5 & 1 & 5 & 2 & 3 & 3 
\end{bmatrix}
}\end{equation}

@@

Finally, we define the \emph{sample mechanism} \(S\) in Equation
(\ref{eq-mat25}), which maps each element of \(S_{IA}\) to every element
of \(S_{JK}\). Each element \(s_{(i,a,h,b,j,k)}\) is a binary value
indicating the presence or absence of data rows in the matrix \(R\)
resulting from the sample mechanism, as in Equation (\ref{eq-mat26}).
Thus, the vector contains \(n\) elements, matching the number of rows in
\(R\), and the sum of its elements represents the total data sample:
\(n^{s} = \binom{n^{s}_{I} \cdot n^{s}_{A}}{2} \cdot n^{s}_{J} \cdot n^{s}_{K}\).
Here, the term \({n^{s}_{I} \cdot n^{s}_{A} \choose 2}\) represents the
binomial coefficient, which quantifies the total number of unique
comparisons possible between every pair of sampled stimuli generated by
each pair of sampled individuals.
\begin{equation}\phantomsection\label{eq-mat25}{
S = \begin{bmatrix}
s_{(1,1,1,2,1,1)} \\
\vdots \\
s_{(1,1,1,2,1,n_{K})} \\
\vdots \\
s_{(i,a,h,b,j,k)} \\
\vdots \\
s_{(n_{I},n_{A}-1,n_{I},n_{A},n_{J},1)} \\
\vdots \\
s_{(n_{I},n_{A}-1,n_{I},n_{A},n_{J},1)}
\end{bmatrix}
}\end{equation}

@@

Note that while the joint distribution in Equation (\ref{eq-mat51})
includes the probability distributions of the sampling and comparison
mechanisms, \(P(S)\) and \(P(C)\), as well as those of the predictor
variables--\(P(X_{IA})\), \(P(X_{I})\), \(P(Z_{JK})\), and
\(P(Z_{J})\)--we omit all of these probabilities from the statistical
model \ref{fig-cj16_stat}. This omission is justified because, while
these distributions contribute to the overall joint distribution of the
data, the variables \(S\), \(C\), \(X_{IA}\), \(X_{I}\), \(Z_{JK}\), and
\(Z_{J}\) are observed and independent of any other variable in the
model. As observed variables, they do not require distributional
assumptions in the same way the idiosyncratic errors do. Furthermore,
their independence follows from the underlying random selection
procedures that govern the variables\footnote{Randomization ensures that
  data--and, by extension, an estimator--satisfies several key
  identification properties, such as common support, no interference,
  and consistency. The most critical property, however, is the
  elimination of confounding. \emph{Confounding} occurs when an external
  variable, such as \(X_{I}\), simultaneously influences both the
  outcome (e.g., \(O_{R}\)) and a variable of interest (e.g., \(S\)),
  resulting in spurious associations between the latter two
  \citep{Everitt_et_al_2010}. Randomization ensure the absence of
  confounding by effectively decoupling the association between the
  variable of interest and any other variable, except for the outcome
  itself. For a more detailed discussion on the benefits of
  randomization, see \citet{Pearl_2009}, \citet{Morgan_et_al_2014},
  \citet{Neal_2020}, and \citet{Hernan_et_al_2025}.}.

@@

Thurstone introduced the Law of Comparative Judgment to measure
psychological traits of stimuli through pairwise comparisons
\citep{Thurstone_1927a, Thurstone_1927b}. In its general form, the
theory models single-judge comparisons across multiple, potentially
correlated stimuli. Each comparison produces a dichotomous outcome
indicating which stimulus the judge perceives as having a higher trait
level. However, Thurstone identified one key challenge with its
measurement model: it required estimating more ``unknown'' parameters
than the number of available pairwise comparisons
\citep{Thurstone_1927b}. To address this issue and facilitate the
theory's practical applicability, he formulated five cases, each
progressively incorporating several simplifying assumptions.

Among these, Case V has become the most widely used model in empirical
CJ research, primarily due to the widespread adoption of the BTL model.
The BTL model mirrors the core assumptions of Case V--namely, equal
discriminal dispersions and zero correlation among stimuli' discriminal
processes--but replaces the processes' normal distribution with the more
mathematically tractable logistic distribution
\citep{Andrich_1978, Bramley_2008}. Although the substitution has a
minimal impact on trait estimation or model interpretation
\citep{vanderLinden_et_al_2017_I, McElreath_2021} and simplifies the
measurement model, the other two core assumptions of the model can still
lead to unreliable and inaccurate trait estimates
\citep{Ackerman_1989, Zimmerman_1994, McElreath_2020, Hoyle_et_al_2023},
especially when researchers assess complex traits or heterogeneous
stimuli
\citep{Thurstone_1927a, Andrich_1978, Bramley_2008, Kelly_et_al_2022}.

Furthermore, because Thurstone primarily aimed to produce a trait's
``coarse scaling'' and allocate the compared stimuli along this
continuum \citep[pp.~269]{Thurstone_1927a}, his theory did not specify
how to use the trait estimates for statistical inference. The CJ
tradition has attempted to fill this gap by separating trait estimation
from hypothesis testing. Specifically, researchers often rely on point
estimates of the traits, such as the BTL scores or their
transformations, for inference. Although this approach simplifies
analysis, the statistical literature cautions that it can introduce bias
and compromise the reliability of the resulting inferences
\citep{McElreath_2020, Kline_et_al_2023, Hoyle_et_al_2023}.

Thus, to address these limitations of Thurstone's Case V and the BTL
model, this study extends Thurstone's general form using a systematic,
integrated approach that combines Causal and Bayesian inference methods.
The approach begins by formulating a conceptual model represented by a
Structural Causal Model (SCM) and a Directed Acyclic Graph (DAG). This
model integrates Thurstone's core theoretical principles, such as the
discriminal processes of stimuli, alongside key CJ experimental design
features, including judges' bias, sampling procedures, and comparison
mechanisms, thereby disentangling the causal processes in the CJ system.

@@

Despite its promise, this study presents several challenges that
researchers must navigate --challenges common to many quantitative
studies. These challenges fall into two main categories. The first
concerns the background knowledge researchers need to apply the proposed
framework effectively. The second involves the increased attention
researchers must now devote to their specific CJ research context.

In terms of background knowledge, researchers face two challenges:
mastering Causal inference and developing proficiency in Bayesian
inference methods. One example illustrating the need to master Causal
inference is our assumption that predictor variables are ``relevant'' to
the research context--interpreting relevance as the inclusion of these
variables in a ``sufficient adjustment set''
\citep{Pearl_2009, Pearl_et_al_2016, Morgan_et_al_2014}. However, we do
not fully explore what this entails or its implications for model
specification, estimation, and inference. As a result, researchers must
engage deeply with these complex ideas--and with others, such as how
SCMs encode functional and probabilistic information--to apply the
framework rigorously. To assist this effort, we provide key references
throughout the study to guide motivated readers toward a deeper
understanding of these foundational concepts\footnote{refer to footnote
  1, 2, and the detailed online document referenced in the Declarations
  section}.

Bayesian inference presents a second challenge. Researchers must acquire
sufficient expertise to translate context-specific conceptual models
into bespoke statistical models that meaningfully address their research
questions. Although such translations are not unique to Bayesian
approaches, we argue that Bayesian methods offer a more accessible
route, as they reduce the need for specialized knowledge in optimization
theory and techniques--often required by frequentist methods--thereby
lowering the technical barriers for many applied researchers.

In terms of the attention researchers must now devote to their specific
CJ research context, researchers face two challenges: ensuring the
conceptual model is a \emph{faithful} representation for the CJ system,
and verifying whether this model can accurately estimate the intended
estimand from real data.

@@

This section introduces fundamental statistical and causal inference
concepts necessary for understanding the core theoretical principles
described in this document. It does not, however, offer a comprehensive
overview of causal inference methods. Readers seeking more in-depth
understanding may wish to explore introductory papers such as
\citet{Pearl_2010}, \citet{Rohrer_2018}, \citet{Pearl_2019}, and
\citet{Cinelli_et_al_2020}. They may also find it helpful to consult
introductory books like \citet{Pearl_et_al_2018}, \citet{Neal_2020}, and
\citet{McElreath_2020}. For more advanced study, readers may refer to
seminal intermediate papers such as \citet{Neyman_et_al_1923},
\citet{Rubin_1974}, \citet{Spirtes_et_al_1991}, and \citet{Sekhon_2009},
as well as books such as \citet{Pearl_2009}, \citet{Morgan_et_al_2014},
and \citet{Hernan_et_al_2025}.
